---
title: "Final Team Project"
author: "Mitchell Fuchs, Lando Schwerdtfeger, Nick Lo, Maria Gonzalez, Tomas Llopart"
date: "2025-11-19"
output:
  html_document:
    toc: true
    toc_float: true
    toc_collapsed: true
    toc_depth: 3
    highlight: tango
    theme: cerulean
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(warning = FALSE)
knitr::opts_chunk$set(message = FALSE)
```

# Step 0 - Why?

# Step 1 & 2 - Load and Clean the Data

# Step 3 - Split the Data

# Step 4 - Build the Models

# Step 5 - Predict the Models

We estimated three Support Vector Machine models using linear, radial, and polynomial kernels.  
The purpose of this module is to evaluate how different kernel functions perform on our scaled feature set.  
To keep the workflow modular and reproducible, all SVM training and evaluation steps are contained in the external script `SVM_Model.R`.

```{r DT_cm}
source("Data/split_data.R")
source("Models/Decision_Tree.R")

# Display results
cm_dt
```


```{r SVM_cm}
source("Data/split_data.R")
source("Models/SVM_Model.R")
best_cm

```

The highest accuracy indicates the best-performing SVM kernel for this dataset.  
Below we display the confusion matrix for the best model, as well as a comparison plot across all kernels.


# Step 6 - Evaluate the Models


# Step 7 - Implement final Model
